---
title: "BioconductorHW.Rmd"
author: "Jose Nieves"
date: "April 7, 2016"
output: html_document
---

#Bioconductor packages

Pick one of the hundreds of bioconductor packages listed on the site that contain analysis or data of interest to you. Find the vignettes for the package, and turn in an example analysis using the package. Try and use markdown to write your example, and upload the .Rmd file. I have linked to the bioconductor package list and the help for markdown in RStudio.

####Download and install package from bioconductor
```{r}
# set working directory
setwd("/Users/jose_nieves/Documents/School/INTD_8065/GitHub.INTD-8065/Homework.INTD-8065")

# install bioconductor package
source("https://bioconductor.org/biocLite.R")
biocLite("caOmicsV")

# install caOmicsV package
library(caOmicsV)

# read vignettes for the caOmicsVR package
browseVignettes("caOmicsV")
```

###Generate a bioMatrix layout image with the build-in demo data
```{r}
data(biomatrixPlotDemoData)

plotBioMatrix(biomatrixPlotDemoData, summaryType="text")
bioMatrixLegend(heatmapNames=c("RNASeq", "miRNASeq"), 
                categoryNames=c("Methyl H", "Methyl L"), 
                binaryNames=c("CN LOSS", "CN Gain"),   
                heatmapMin=-3, heatmapMax=3, colorType="BlueWhiteRed")
```

###BioNetCircos layout image with the build-in demo data
```{r}
data(bionetPlotDemoData)

plotBioNetCircos(bionetPlotDemoData)
dataNames <- c("Tissue Type", "RNASeq", "miRNASeq", "Methylation", "CNV")
bioNetLegend(dataNames, heatmapMin=-3, heatmapMax=3)
```

#####making plot data set (list of function's arguments)
```{r}
args(getPlotDataSet) 
```

###Plot biomatrix layout
```{r}
#####1. Demo data
data("biomatrixPlotDemoData")
dataSet <- biomatrixPlotDemoData
names(dataSet)

#####2. initialize biomatrix layout
numOfGenes <- length(dataSet$geneNames)
numOfSamples <- length(dataSet$sampleNames) 
numOfPhenotypes <- nrow(dataSet$sampleInfo)-1

numOfHeatmap <- length(dataSet$heatmapData)
numOfSummary <- length(dataSet$summaryData)
phenotypes   <- rownames(dataSet$sampleInfo)[-1]

sampleHeight <- 0.4
sampleWidth <- 0.1 
samplePadding <- 0.025
geneNameWidth <- 1
sampleNameHeight <- 2.5
remarkWidth <- 2
summaryWidth <- 1
rowPadding <- 0.1

initializeBioMatrixPlot(numOfGenes, numOfSamples, numOfPhenotypes, 
                        sampleHeight, sampleWidth, samplePadding, rowPadding,geneNameWidth, remarkWidth, summaryWidth, sampleNameHeight)
caOmicsVColors <- getCaOmicsVColors()

par(cex=0.75)
showBioMatrixPlotLayout(dataSet$geneNames,dataSet$sampleNames, phenotypes)

#####3. plot tissue types on phenotype area
head(dataSet$sampleInfo)[,1:3]
rowIndex <- 2

sampleGroup <- as.character(dataSet$sampleInfo[rowIndex,])
sampleTypes <- unique(sampleGroup)
sampleColors <- rep("blue", length(sampleGroup))
sampleColors[grep("Tumor", sampleGroup)] <- "red"

rowNumber <- 1
areaName <- "phenotype"

plotBioMatrixSampleData(rowNumber, areaName, sampleColors)

geneLabelX <- getBioMatrixGeneLabelWidth()
maxAreaX <- getBioMatrixDataAreaWidth()
legendH <- getBioMatrixLegendHeight()
plotAreaH <- getBioMatrixPlotAreaHeigth()
sampleH<- getBioMatrixSampleHeight()

sampleLegendX <- geneLabelX + maxAreaX
sampleLegendY <- plotAreaH + legendH - length(sampleTypes)*sampleH
colors <- c("blue", "red")
legend(sampleLegendX, sampleLegendY, legend=sampleTypes, 
       fill=colors,  bty="n", xjust=0)

#####4. heatmap plot
heatmapData <- as.matrix(dataSet$heatmapData[[1]][,])
plotBioMatrixHeatmap(heatmapData, maxValue=3, minValue=-3)

heatmapData <- as.matrix(dataSet$heatmapData[[2]][,])
plotBioMatrixHeatmap(heatmapData, topAdjust=sampleH/2,  
                     maxValue=3, minValue=-3)

secondNames <- as.character(dataSet$secondGeneNames)
textColors <- rep(caOmicsVColors[3], length(secondNames))
plotBioMatrixRowNames(secondNames, "omicsData", textColors, 
                      side="right", skipPlotColumns=0)

#####5. draw outlines for each samples to show methylation status
categoryData <- dataSet$categoryData[[1]]
totalCategory <- length(unique(as.numeric(dataSet$categoryData[[1]])))

plotColors <- rev(getCaOmicsVColors())
plotBioMatrixCategoryData(categoryData, areaName="omicsData", 
                          sampleColors=plotColors[1:totalCategory])

#####6. binary data plot
binaryData <- dataSet$binaryData[[1]]
plotBioMatrixBinaryData(binaryData, sampleColor=caOmicsVColors[4])

binaryData <- dataSet$binaryData[[2]]
plotBioMatrixBinaryData(binaryData, sampleColor=caOmicsVColors[3])

#####7. plot summary data on right side of plot area
summaryData  <- dataSet$summaryInfo[[1]][, 2]
summaryTitle <- colnames(dataSet$summaryInfo[[1]])[2]

remarkWidth <- getBioMatrixRemarkWidth()
sampleWidth <- getBioMatrixSampleWidth()
col2skip <- remarkWidth/2/sampleWidth + 2

plotBioMatrixRowNames(summaryTitle, areaName="phenotype", 
                      colors="black", side="right", skipPlotColumns=col2skip)

plotBioMatrixRowNames(summaryData, "omicsData", 
                      colors=caOmicsVColors[3], side="right", 
                      skipPlotColumns=col2skip)

#####8. add legend
bioMatrixLegend(heatmapNames=c("RNASeq", "miRNASeq"), 
                categoryNames=c("Methyl H", "Methyl L"), 
                binaryNames=c("CN LOSS", "CN Gain"),   
                heatmapMin=-3, heatmapMax=3, 
                colorType="BlueWhiteRed")
dev.off()
```

###plot bionetcircos layout
```{r}
#####1. Demo data
data(bionetPlotDemoData)
dataSet <- bionetPlotDemoData

sampleNames  <- dataSet$sampleNames
geneNames    <- dataSet$geneNames
numOfSamples <- length(sampleNames)

numOfSampleInfo <- nrow(dataSet$sampleInfo) - 1
numOfSummary <- ifelse(dataSet$summaryByRow, 0, col(dataSet$summaryInfo)-1)
numOfHeatmap <- length(dataSet$heatmapData)
numOfCategory <- length(dataSet$categoryData)
numOfBinary <- length(dataSet$binaryData)

expr <- dataSet$heatmapData[[1]]
bioNet <- bc3net(expr) 

# 2. initialize bionetcircos layout
widthOfSample    <- 100
widthBetweenNode <- 3
lengthOfRadius   <- 10

dataNum <- sum(numOfSampleInfo, numOfSummary, numOfHeatmap, 
               numOfCategory, numOfBinary)
trackheight <- 1.5
widthOfPlotArea  <- dataNum*2*trackheight

initializeBioNetCircos(bioNet, numOfSamples, widthOfSample, 
                       lengthOfRadius, widthBetweenNode,widthOfPlotArea)
caOmicsVColors <- getCaOmicsVColors()
supportedType <- getCaOmicsVPlotTypes()

par(cex=0.75)
showBioNetNodesLayout()

#####3. label each node
par(cex=0.6)
onTop <- c(14, 15, 16, 9, 7, 20, 8, 24, 10, 25)
labelBioNetNodeNames(nodeList=onTop,labelColor="blue", 
                     labelLocation="top", labelOffset = 0.7)

onBottom <- c(26, 22, 23, 18, 19, 3, 5)
labelBioNetNodeNames(nodeList=onBottom,labelColor="black", 
                     labelLocation="bottom", labelOffset = 0.7)

onLeft <- c(2, 11, 21, 17)
labelBioNetNodeNames(nodeList=onLeft,labelColor="red", 
                     labelLocation="left", labelOffset = 0.7)

onRight <- c(13, 12, 4, 1, 6)
labelBioNetNodeNames(nodeList=onRight,labelColor="brown", 
                     labelLocation="right", labelOffset = 0.7)
eraseBioNetNode() # once nodes are labelled plot area erased

#####4. plot each data set
inner <- lengthOfRadius/2
outer <- inner +  trackheight

# plot tissue type for each node
groupInfo <- as.character(dataSet$sampleInfo[2, ])
sampleColors <- rep("blue", numOfSamples);
sampleColors[grep("Tumor", groupInfo)] <- "red"

plotType=supportedType[1]
groupInfo <- matrix(groupInfo, nrow=1)
bioNetCircosPlot(dataValues=groupInfo, 
                 plotType, outer, inner, sampleColors)

inner <- outer + 0.5
outer <- inner +  trackheight  

# heatmap plot for each node
exprData <- dataSet$heatmapData[[1]]
plotType <- supportedType[4] 
bioNetCircosPlot(exprData, plotType, outer, inner, 
                 plotColors="BlueWhiteRed", maxValue=3, minValue=-3)

inner <- outer + 0.5 
outer <- inner +  trackheight

# category data plot for each node
categoryData <- dataSet$categoryData[[1]]
plotType <- supportedType[2]
bioNetCircosPlot(categoryData, plotType, 
                 outer, inner, plotColors="red")
inner <- outer + 0.5
outer <- inner +  trackheight

# binary data plot for each node
binaryData <- dataSet$binaryData[[1]]
plotType <- supportedType[3]
plotColors <- rep(caOmicsVColors[1], ncol(binaryData))
bioNetCircosPlot(binaryData, plotType, 
                 outer, inner, plotColors)

inner <- outer + 0.5
outer <- inner +  trackheight 

# link samples on a node
outer <- 2.5
bioNetGraph <- getBioNetGraph()
nodeIndex <- which(V(bioNetGraph)$name=="PLVAP")

fromSample <- 10 
toSample <- 50 
plotColors <- "red"
linkBioNetSamples(nodeIndex, fromSample, 
                  toSample, outer, plotColors)

fromSample <- 40 
toSample <- 20 
plotColors <- "blue"
linkBioNetSamples(nodeIndex, fromSample, 
                  toSample, outer, plotColors)

#####5. add legend
dataNames <- c("Tissue Type", "RNASeq", "Methylation", "CNV")
bioNetLegend(dataNames, heatmapMin=-3, heatmapMax=3)
dev.off()
```

