---
title: "Correlation and Regression"
author: "M. E. Perez"
date: "February 25, 2016"
output: html_document
---

## Correlation

Population definition:

$$ \rho_{Y_1Y_2} =\frac{
E\left[(Y_1-E(Y_1))(Y_2-E(Y_2))\right]}{\sigma_{Y_1}\sigma_{Y_2}}
$$

Sample estimator:
$$ r_{Y_1Y_2} = \frac{\frac{1}{n-1} \sum_{i=1}^n(y_{i1}-\bar{y}_1)(y_{i2}-\bar{y}_2)}{s_{Y_1}s_{Y_2}}$$

```{r}
data(trees)
pairs(trees)
```

The graph shows a strong association between Volume and Girth, and
some degree of association between Height and Volume. The
explanatory variables also seem to be associated.

```{r}
cor(trees)
```


## Simple Regression Model

$$y_i  =   \beta_0+\beta_1 x_i + \varepsilon_i, \; \;
 i=1,\ldots,n$$
 
where $\varepsilon_i \sim  N(0,\sigma^2)$

```{r}
trees.mod1 = lm(Volume~Girth, data=trees)
names(trees.mod1)
coef(trees.mod1)
```

## Tests of Hypothesis on the Parameters of the Model

$$ H_0: \; \; \beta_i = 0 \; \;\mbox{   vs.   } \beta_i \neq 0$$

When $H_0$ is true 

$$ t=\frac{\hat{\beta_i}}{\mbox{Standard Error}(\hat{\beta}_i)}\sim t_{n-2}$$
(null distribution)


```{r}
summary(trees.mod1)
```

$\hat{\beta}_0$ is negative, and so it can't be
interpreted as the volume of timber when the girth approaches
zero!

We reject  $H_0: \beta_1=0$ (p-value $< 2 \times 10^{-16}$). 
The volume of timber can be predicted using the girth as explanatory variable.

## F-test and Multiple determination coefficient $R^2$

The last line of the summary of the model contains the test statistic and the p-value for comparing our model with the model $\mbox{Volume}_i=\beta_0 + \varepsilon_i$. In our example, $F_{1,29}= 419.4$ and $\mbox{p-value}<2.2 \times 10^{-16}$ (lower than the smallest value that can be represented by the computer).

Also in the summary of the model we can find the value of the multiple determination coefficient $R^2$ (a measure of goodness of fit).In our example, $R^2 = 0.9353$, indicating that the model fits well the data.

## Residual plots

```{r}
par(mfrow=c(2,2))
plot(trees.mod1,pch=16)
```

* The residuals vs fitted values plot has a curved form (positive residuals in the extremes, negative residuals at the center). This suggests that we should adjust a different model with higher order terms.

* Observation 31 is an influential observation, and it seems to be an outlier with respect to the model (an observation with an unusually large residual)

```{r}
par(mfrow=c(1,1))
with(trees,plot(Girth,Volume,pch=16))
with(trees,text(Girth[31],Volume[31]-2,labels="31",cex=0.8))
```

# Multiple linear regression

```{r}
trees.mod2=lm(Volume~Girth+I(Girth^2)+Height,data=trees)
summary(trees.mod2)
```

It is not advisable to remove more than one variable at a time using the t-test, as the t-test only allows to compare models differing in one variable.

## Analysis of variance

```{r}
trees.mod0 = lm(Volume ~ 1, data=trees)
summary(trees.mod0)
anova(trees.mod0,trees.mod1)
```

The F-statistic and the p-value are the ones in the summary of the model.

## Confidence intervals and prediction intervals

We will work with a "mean" tree.

```{r}
with(trees,c(mean(Girth),mean(Height)))
```

Confidence interval for a "mean" tree.

```{r}
predict(trees.mod1,data.frame(Girth=13.25,Height=76),interval="confidence")
```

Prediction interval  for a "mean" tree.

```{r}
predict(trees.mod1,data.frame(Girth=13.25,Height=76),interval="prediction")
```

