---
title: "Homework.5"
author: "Jose Nieves"
date: "February 26, 2016"
output: html_document
---
#INTD 8065 Data Analysis for Cancer Research
#Homework #5

1. Run the t test for the comparison of control diet         and high fat diet presented in page 51-52 of              Irizarry and Love (2015) assuming equality                ofvariances. 
        
Compare your result with that of the analysis obtained in the book.

```{r}
library(dplyr)
library(downloader)
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/
        extdata/femaleMiceWeights.csv"
filename <- "femaleMiceWeights.csv"
if (!file.exists(filename)) download(url, destfile=filename)
femaleMiceWeights <-read.csv(filename)

control <- filter(femaleMiceWeights,Diet=="chow") %>% select(Bodyweight) %>% unlist
treatment <- filter(femaleMiceWeights,Diet=="hf") %>% select(Bodyweight) %>% unlist

t.test(treatment, control, var.equal = TRUE)
t.test(treatment, control)
```

In what follows, we will use library (pwr) mentioned in Irizarry and Love (2015),  page 64.

2. Download and install the library.

```{r}
library(pwr)
```

3. Use the function power.t.test in library pwr for          calculating the power obtained by simulation in page 64    by Irizarry and Love for N=12 and alpha = 0.05. 
        
Use delta as the real difference between the population means. Assume the standard deviations of the two populations are equal and calculate it as
        
sd <- sqrt(((n1-1)*var(hfPopulation)+(n2-1)*var(controlPopulation))/(n1+n2))
        
where n1 and n2 are the sizes of hfPopulation and controlPopulation, respectively
        
```{r}
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/
        extdata/mice_pheno.csv"
filename <- "mice_pheno.csv"
if (!file.exists(filename)) download(url, destfile=filename)
mice_pheno <- read.csv(filename)

controlPopulation <- filter(mice_pheno,Sex == "F" & Diet == "chow") %>%
                select(Bodyweight) %>% unlist
hfPopulation <- filter(mice_pheno,Sex == "F" & Diet == "hf") %>% 
                select(Bodyweight) %>% unlist

N <- 12
alpha <- 0.05
B <- 2000
n1 <- nrow(filter(mice_pheno,Sex == "F" & Diet == "hf"))
n2 <- nrow(filter(mice_pheno,Sex == "F" & Diet == "chow"))
sd <- sqrt(((n1-1)*var(hfPopulation)+(n2-1)*var(controlPopulation))/(n1+n2))

power.t <- power.t.test(12, delta = TRUE, sd, sig.level = alpha, power = NULL)
print(power.t$power)
```        

4. Use now  power.t.test for calculating the power for       N=5, 10, 15, 20, 25,30, 35, 40, 45, 50 (this is the Ns    sequence created for the graph in page 66). Overimpose    your results to the graph generated in the book using     red lines and dots.

Hint: You can pass a vector of N’s to the function, and         extract the power values from the object you created       using object$power. For overimposing your results,        you can use the command line with the option              type=”b”.
```{r, echo=FALSE}
Ns <- seq(5, 50, 5)
power.Ns <- power.t.test(n = Ns, delta = TRUE, sd, sig.level = alpha, power = NULL)
plot(Ns,power.Ns$power,
     main = "Power plotted against sample size",
     type = "b",
     las = 1,
     ylab = "Power",
     col = 4)
```      

5. Repeat part 3. for the graph of power for different       values of alphas shown in page 67.

```{r, echo=FALSE}
alphas <- c(0.1,0.05,0.01,0.001,0.0001) 
power.alphas <- power.t.test(n = 30, delta = T, sd, sig.level = alphas, power = NULL)
plot(alphas, power.alphas$power,
     main = "Power plotted against cutt-off",
     xlab = "alpha",
     ylab = "Power",
     type = "b",
     log = "x",
     las = 1)
```      

6. In this setting, use power.t.test for calculating the     sample sizes for obtaining powers of 0.25, 0.5, 0.75      and 0.85. Use a = 0.05 and a = 0.01. Comment your         results.

Made wrapper for a for loop over \code{power.t.test}.

```{r}
Power.t.test <- function(n=NULL, delta=NULL, sd=1, sig.level=0.05,
                power=NULL, type="two.sample",
                alternative = "two.sided",
                strict=FALSE){
        
        if (sum(sapply(list(n, delta, power), is.null)) != 1)
                stop("Exactly one of 'n', 'delta', and 'power' may be NULL")
        
        extract <- names(which(sapply(list(n=n, delta=delta, power=power), is.null)))
        
        Power <- expand.grid(n = if (is.null(n)) NA else n,
                        delta = if (is.null(delta)) NA else delta,
                        sd = sd,
                        sig.level = sig.level,
                        power = if (is.null(power)) NA else power,
                        type = type,
                        alternative = alternative,
                        strict = strict,
                        stringsAsFactors=FALSE)
        for (i in 1:nrow(Power)){
                Power[[extract]][i] <- with(Power, power.t.test(n = if (is.na(n[i])) NULL else n[i], 
                                delta = if (is.na(delta[i])) NULL else delta[i], 
                                sd = sd[i],
                                sig.level = sig.level[i], 
                                power = if (is.na(power[i])) NULL else power[i],
                                type = type[i],
                                alternative = alternative[i],
                                strict = strict[i]))[[extract]]
        }
        
        return(Power)
        
}   
```

vectors for powers and alphas

```{R}
powers.2 <- c(0.25,0.5,0.75,0.85)
alphas.2 <- c(0.05, 0.01)
```

power.t.test to find n using a = 0.05, 0.01 & powers 0.25,0.5,0.75,0.85

```{R}
n.pwr.t <- Power.t.test(n = NULL, delta = T, sd = sd, sig.level = alphas.2,
                power = powers.2)
keeps <- c("n", "sig.level", "power")
n.pwr.t[keeps]
```

